Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-22 WSSSPE1_report/WSSSPE1_summary_report.tex
adding some text

Another paradigm in utility software is the software delivered as service over the web.

\subsection{Scientific Software}
Special purpose software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind. Software projects pertaining to specific scientific domain often tend to be in a niche and the user community tends to be small to medium. Numerical accuracy and algorithmic optimization are some of the paramount requirements of most scientific software. Advantages of a smaller codebase and requirements include stability, ease of installation, configurability and integration into larger systems. While the software can stay stable and require relatively low maintenance, the responsibility is often on the shoulders of a very few developers. Community participation in the development tends to be low. Development is linear and simplistic with a limited scope to follow software best practices. Domain science oriented software developed to solve  specific problems, e.g., \emph{DUNE}, \emph{PETSc}, \emph{MINRES-QLP}, \emph{FASTMath}. One exception is \emph{Kitware}.

>>>>

Another paradigm in utility software is the software delivered as service over
the web. With increasing popularity of cloud-based computational environments,
many users are leaning towards tools used as service. \emph{GitHub} can be
arguably considered one such tool, catering to collaborative development. For
scientific users \emph{Globus} based tools are case of service based utility
software discussed during the workshop. The data movement, authentication and
sharing services offered by Globus can be easily used over the web by
collaborating researcher.

\subsection{Scientific Software}
Special purpose software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind. Software projects
pertaining to specific scientific domain often tend to be in a niche and the
user community tends to be small to medium. Numerical accuracy and algorithmic
optimization are some of the paramount requirements of most scientific
software. Advantages of a smaller codebase and requirements include stability,
ease of installation, configurability and integration into larger systems.
While the software can stay stable and require relatively low maintenance, the
responsibility is often on the shoulders of a very few developers. Community
participation in the development tends to be low. Development is linear and
simplistic with a limited scope to follow software best practices. Domain
science oriented software developed to solve  specific problems, e.g.,
\emph{DUNE}, \emph{PETSc}, \emph{MINRES-QLP}, \emph{FASTMath}. One exception is
\emph{Kitware}. Sustainability of such software is often a significant
challenge. Many submissions report the software considered as a `byproduct` of
actual research. Others contend that the software was not the main funded part
of their research.
